A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2021-01-18 cusp_j/paper/submit/paper.tex
Publishable version

Coming now to the behaviour of $u(r_{12};\mu)$ as a function of the parameter $\mu$, we report in Figure \ref{fig_j_mu} its dependence in $r_{12}$ for a set of values of $\mu$, and we compare it with the FROGG. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/small_mu_j.pdf}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/small_mu_exp_j.pdf}\\
        \caption{Shape of $u(r_{12};\mu)$ (left) and of $\text{exp}\big(u(r_{12};\mu) \big) $ (right) for different values of $\mu$. 
We also compare the FROGG (see Eq. \eqref{eq_frogg}) and the value of $\mu = 0.865$ which equals the FROGG at $r_{12}=0$ (see Eq. \eqref{mu_ten_no_1}). }
 \label{fig_j_mu}
\end{figure*}
From Figure\ref{fig_j_mu}, one can see that: i) the larger the $\mu$, the more short-range is $u(r_{12};\mu)$, ii) the larger the $\mu$, the closer $u(r_{12};\mu)$ is from 0 and therefore the less impact the correlating factor has on the wave function, iii) the smaller the $\mu$, the deeper and broader is the hole imposed by $e^{u(r_{12},\mu)}$.

@@

As the present study is devoted to two-electron systems, the physics of $\tilde{H}[\mu]$ can be qualitatively understood by looking at $\tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$, which is illustrated in Fig. \ref{fig_wee_mu} for different values of $\mu$ and compared with the usual long-range interaction of the RS-DFT framework. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/w_ee_small.pdf}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/w_ee_erf.pdf}\\
        \caption{Shape of $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ (left) and comparison with $\frac{\text{erf}(\mu r_{12}}{r_{12}}$ (right) for different values of $\mu$.}
 \label{fig_wee_mu}
\end{figure*}
From  Fig. \ref{fig_wee_mu} we can see that the scalar effective potential $ \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$ 
is non divergent, increases when $\mu$ increases and globally resembles the long-range interaction used in RS-DFT at large $r_{12}$. 
Nevertheless, as observed from Fig. \ref{fig_wee_mu}, it is significantly different 
from the long-range interaction at small $r_{12}$, 
and it is not monotonic and attractive (\textit{i.e.} $\lim_{r_{12} \rightarrow 0} \deriv{}{r_{12}}{} \tilde{\mathcal{W}}_{ee}(r_{12},\mu) >0 $). 
This difference with respect to the long-range interaction can be understood as the equation obtained to derive the Jastrow factor only takes into account the first-order derivative of $u(r_{12},\mu)$ (see Eqs. \eqref{def_j_00} and \eqref{def_j_0}). 
If one desires to impose that $\tilde{\mathcal{W}}_{ee}(r_{12},\mu) = \frac{\text{erf}(\mu r_{12})}{r_{12}}$, one would have to solve the non-linear differential equation 
\begin{equation}
 2 \deriv{u(r_{12})}{r_{12}}{} + r_{12} \bigg( \deriv{u(r_{12})}{r_{12}}{2} + \bigg[ \deriv{u(r_{12})}{r_{12}}{} \bigg]^2\bigg) = 1 - \text{erf}(\mu r_{12}), 
\end{equation}
whose solution is unknown to the best of the author knowledge.

@@

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/w_ee_frog.pdf}
        \caption{Shape of $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ for $\mu=\mfrogg$ and comparison with the effective potential $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ obtained with the FROGG (see Eq. \eqref{eq:_wee_frogg}).}
 \label{fig_wee_frogg}
\end{figure*}

@@

From the study of the ground state of the helium atom, one can conclude that i) the use of a simple Jastrow factor such as $u(r_{12};\mu)$ within the transcorrelated framework can strongly improve the speed of convergence of the total energy of the helium atom with a quite wide range of values of $\mu$, 
ii) the results obtained with $\mu=\mfrogg$ reproduce essentially that obtained with the FROGG, 
iii) for $\mu \in[ 0.35,0.7]$ the results obtained are always better than that obtained with the FROGG. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/He/Energy/He_E_conv_basis_small_mu.pdf}
        \includegraphics[width=0.45\linewidth]{plots/He/Energy/He_E_conv_basis_large_mu.pdf}\\
        \caption{Convergence of the ground state eigenvalue of $\tilde{H}[\mu]$ with respect to the AVXZ basis set series (X=D,T,Q,5) for different values of $\mu$. 
We also compare with the results of the TC obtained with the $u(r_{12}) = u^{\text{TN}}(r_{12})$ (see Eq. \eqref{eq_frogg}), referred here after as FROGG. }
 \label{fig_conv_e_mu}
\end{figure*}

@@

We report in Figs \ref{fig:mu_0.3_dz_3} and \ref{fig:mu_1.0_dz_3} the plot of the right eigenvector $\phimub(\br{}_1,\br{}_2)$, estimated exact wave function $\psimub(\br{}_1,\br{}_2)$ for $\mu=0.3$ and $\mu=1.0$ (respectively) and compared to the exact wave function $\psiex(\br{}_1,\br{}_2)$ and FCI wave function for $r=0.5\,a.u.$. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/He/cusp/He_mu_0_3_cusp_avdz_avtz_3.pdf}
        \includegraphics[width=0.45\linewidth]{plots/He/cusp/He_mu_0_3_cusp_avqz_av5z_3.pdf}\\
        \caption{
        Helium atom, radius $r=0.5$ a.u.: approximation of the exact wave function $\psimub(\br{}_1,\br{}_2)$ and right eigenvector $\phimub(\br{}_1,\br{}_2)$ for $\mu=0.3$ in the AVDZ and AVTZ basis sets (left) and AVQZ and AV5Z (right). Comparison with the FCI wave function in the same basis sets and the estimated exact wave function $\psiex(\br{}_1,\br{}_2)$.  }
 \label{fig:mu_0.3_dz_3}
\end{figure*}

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/He/cusp/He_mu_1_0_cusp_avdz_avtz_3.pdf}
        \includegraphics[width=0.45\linewidth]{plots/He/cusp/He_mu_1_0_cusp_avqz_av5z_3.pdf}\\
        \caption{
        Helium atom, radius $r=0.5$ a.u.: approximation of the exact wave function $\psimub(\br{}_1,\br{}_2)$ and right eigenvector $\phimub(\br{}_1,\br{}_2)$ for $\mu=1.0$ in the AVDZ and AVTZ basis sets (left) and AVQZ and AV5Z (right). Comparison with the FCI wave function in the same basis sets and the estimated exact wave function $\psiex(\br{}_1,\br{}_2)$.  }
 \label{fig:mu_1.0_dz_3}
\end{figure*}

@@

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/jastrow/w_ee_zoom.pdf}
        \caption{
        Effective potential $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ for $\mu=0.3$ and $\mu=1.0$, compared to the usual $1/r_{12}$ Coulomb interaction.}
 \label{fig_wee_compare}
\end{figure*}

@@

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{plots/H2/H2_de_fci_frogg.pdf}
        \includegraphics[width=0.45\linewidth]{plots/H2/H2_de_fci_mu_ten_no.pdf}\\
        \includegraphics[width=0.45\linewidth]{plots/H2/H2_de_fci_lda.pdf}
        \includegraphics[width=0.45\linewidth]{plots/H2/H2_de_fci_rsc.pdf}\\
        \includegraphics[width=0.45\linewidth]{plots/H2/H2_de_fci_mu_0_5.pdf}
        \includegraphics[width=0.45\linewidth]{plots/H2/H2_de_fci_mu_0_3.pdf}
        \caption{
        H$_2$ molecule: Error with respect to the estimated exact ground state energy for the aug-cc-pVDZ and aug-cc-pVTZ basis sets (AVDZ, AVTZ, respectively) as a function of the inter nuclear distance for different values of $\mu$, and comparison with FCI. R$_{eq}$ refers to the equilibrium distance of 1.401 a.u., and the estimated exact non relativistic potential energy curve was taken from Ref. \onlinecite{LieCle-JCP-74a}.  
}
 \label{fig:H2}
\end{figure*}

>>>>

Coming now to the behaviour of $u(r_{12};\mu)$ as a function of the parameter $\mu$, we report in Figure \ref{fig_j_mu} its dependence in $r_{12}$ for a set of values of $\mu$, and we compare it with the FROGG. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig1a.pdf}
        \includegraphics[width=0.45\linewidth]{Fig1b.pdf}\\
        \caption{Shape of $u(r_{12};\mu)$ (left) and of $\text{exp}\big(u(r_{12};\mu) \big) $ (right) for different values of $\mu$. 
We also compare the FROGG (see Eq. \eqref{eq_frogg}) and the value of $\mu = 0.865$ which equals the FROGG at $r_{12}=0$ (see Eq. \eqref{mu_ten_no_1}). }
 \label{fig_j_mu}
\end{figure*}
From Figure\ref{fig_j_mu}, one can see that: i) the larger the $\mu$, the more short-range is $u(r_{12};\mu)$, ii) the larger the $\mu$, the closer $u(r_{12};\mu)$ is from 0 and therefore the less impact the correlating factor has on the wave function, iii) the smaller the $\mu$, the deeper and broader is the hole imposed by $e^{u(r_{12},\mu)}$.

@@

As the present study is devoted to two-electron systems, the physics of $\tilde{H}[\mu]$ can be qualitatively understood by looking at $\tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$, which is illustrated in Fig. \ref{fig_wee_mu} for different values of $\mu$ and compared with the usual long-range interaction of the RS-DFT framework. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig2a.pdf}
        \includegraphics[width=0.45\linewidth]{Fig2b.pdf}\\
        \caption{Shape of $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ (left) and comparison with $\frac{\text{erf}(\mu r_{12}}{r_{12}}$ (right) for different values of $\mu$.}
 \label{fig_wee_mu}
\end{figure*}
From  Fig. \ref{fig_wee_mu} we can see that the scalar effective potential $ \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$ 
is non divergent, increases when $\mu$ increases and globally resembles the long-range interaction used in RS-DFT at large $r_{12}$. 
Nevertheless, as observed from Fig. \ref{fig_wee_mu}, it is significantly different 
from the long-range interaction at small $r_{12}$, 
and it is not monotonic and attractive (\textit{i.e.} $\lim_{r_{12} \rightarrow 0} \deriv{}{r_{12}}{} \tilde{\mathcal{W}}_{ee}(r_{12},\mu) >0 $). 
This difference with respect to the long-range interaction can be understood as the equation obtained to derive the Jastrow factor only takes into account the first-order derivative of $u(r_{12},\mu)$ (see Eqs. \eqref{def_j_00} and \eqref{def_j_0}). 
If one desires to impose that $\tilde{\mathcal{W}}_{ee}(r_{12},\mu) = \frac{\text{erf}(\mu r_{12})}{r_{12}}$, one would have to solve the non-linear differential equation 
\begin{equation}
 2 \deriv{u(r_{12})}{r_{12}}{} + r_{12} \bigg( \deriv{u(r_{12})}{r_{12}}{2} + \bigg[ \deriv{u(r_{12})}{r_{12}}{} \bigg]^2\bigg) = 1 - \text{erf}(\mu r_{12}), 
\end{equation}
whose solution is unknown to the best of the author knowledge.

@@

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig3.pdf}
        \caption{Shape of $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ for $\mu=\mfrogg$ and comparison with the effective potential $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ obtained with the FROGG (see Eq. \eqref{eq:_wee_frogg}).}
 \label{fig_wee_frogg}
\end{figure*}

@@

From the study of the ground state of the helium atom, one can conclude that i) the use of a simple Jastrow factor such as $u(r_{12};\mu)$ within the transcorrelated framework can strongly improve the speed of convergence of the total energy of the helium atom with a quite wide range of values of $\mu$, 
ii) the results obtained with $\mu=\mfrogg$ reproduce essentially that obtained with the FROGG, 
iii) for $\mu \in[ 0.35,0.7]$ the results obtained are always better than that obtained with the FROGG. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig4a.pdf}
        \includegraphics[width=0.45\linewidth]{Fig4b.pdf}\\
        \caption{Convergence of the ground state eigenvalue of $\tilde{H}[\mu]$ with respect to the AVXZ basis set series (X=D,T,Q,5) for different values of $\mu$. 
We also compare with the results of the TC obtained with the $u(r_{12}) = u^{\text{TN}}(r_{12})$ (see Eq. \eqref{eq_frogg}), referred here after as FROGG. }
 \label{fig_conv_e_mu}
\end{figure*}

@@

We report in Figs \ref{fig:mu_0.3_dz_3} and \ref{fig:mu_1.0_dz_3} the plot of the right eigenvector $\phimub(\br{}_1,\br{}_2)$, estimated exact wave function $\psimub(\br{}_1,\br{}_2)$ for $\mu=0.3$ and $\mu=1.0$ (respectively) and compared to the exact wave function $\psiex(\br{}_1,\br{}_2)$ and FCI wave function for $r=0.5\,a.u.$. 
\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig5a.pdf}
        \includegraphics[width=0.45\linewidth]{Fig5b.pdf}\\
        \caption{
        Helium atom, radius $r=0.5$ a.u.: approximation of the exact wave function $\psimub(\br{}_1,\br{}_2)$ and right eigenvector $\phimub(\br{}_1,\br{}_2)$ for $\mu=0.3$ in the AVDZ and AVTZ basis sets (left) and AVQZ and AV5Z (right). Comparison with the FCI wave function in the same basis sets and the estimated exact wave function $\psiex(\br{}_1,\br{}_2)$.  }
 \label{fig:mu_0.3_dz_3}
\end{figure*}

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig6a.pdf}
        \includegraphics[width=0.45\linewidth]{Fig6b.pdf}\\
        \caption{
        Helium atom, radius $r=0.5$ a.u.: approximation of the exact wave function $\psimub(\br{}_1,\br{}_2)$ and right eigenvector $\phimub(\br{}_1,\br{}_2)$ for $\mu=1.0$ in the AVDZ and AVTZ basis sets (left) and AVQZ and AV5Z (right). Comparison with the FCI wave function in the same basis sets and the estimated exact wave function $\psiex(\br{}_1,\br{}_2)$.  }
 \label{fig:mu_1.0_dz_3}
\end{figure*}

@@

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig7.pdf}
        \caption{
        Effective potential $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ for $\mu=0.3$ and $\mu=1.0$, compared to the usual $1/r_{12}$ Coulomb interaction.}
 \label{fig_wee_compare}
\end{figure*}

@@

\begin{figure*}
        \includegraphics[width=0.45\linewidth]{Fig8a.pdf}
        \includegraphics[width=0.45\linewidth]{Fig8b.pdf}\\
        \includegraphics[width=0.45\linewidth]{Fig8c.pdf}
        \includegraphics[width=0.45\linewidth]{Fig8d.pdf}\\
        \includegraphics[width=0.45\linewidth]{Fig8e.pdf}
        \includegraphics[width=0.45\linewidth]{Fig8f.pdf}
        \caption{
        H$_2$ molecule: Error with respect to the estimated exact ground state energy for the aug-cc-pVDZ and aug-cc-pVTZ basis sets (AVDZ, AVTZ, respectively) as a function of the inter nuclear distance for different values of $\mu$, and comparison with FCI. R$_{eq}$ refers to the equilibrium distance of 1.401 a.u., and the estimated exact non relativistic potential energy curve was taken from Ref. \onlinecite{LieCle-JCP-74a}.  
}
 \label{fig:H2}
\end{figure*}
